---
title: 'Statistical report: Phenotyping heart failure patients with iron deficiency'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SCREAM2

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart")
```

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`.  

The median age (IQR) is `r rsdata %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

## Created variables 

```{r npr}
default_kable(metaout, caption = "Comorbidities and outcomes from VAL/KON")
```

The labvalues from SCREAM are selected within 90 days prior to index in SwedeHF, 
where the one closest to index is chosen. If there are two tests of the same lab 
from the same date the mean is calculated. The ferritin, transferrin and 
hb value must be from the same date, the other labs can be from different dates (within the 90 days).

The evaluation of ID/A at 6 months, 1 and 3 years is including patients with at least one available 
lab of ferritin, transferrin and hb at the same date at 6 months after index +/- 3 months, 1 years after index +/- 3 months 
and 3 years after index +/- 3 months. The lab values closest to the respective time points is chosen. 

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1aid}. 

## Baseline characteristics

CHECK VARIABLES IN TABLE AND MODEL.

```{r, child = "../src/tab1.Rmd"}

```

## Follow-up ID/A

```{r, child = "../src/tab_follow.Rmd"}

```

\clearpage

## Predictors of ID at index

The association between baseline covariates and ID was investigated using 
logistic regression partly in the overall cohort and partly separately in EF subgroups. 

A p-value for interaction between EF and the respective variable (with the other variables in the model) is also presented. 
(I did not present the OR ect from the interaction model because it seems 
a bit untransparant when I thought about it, then the ORs would all be from separate models...). 

DO YOU PERHAPS WANT TO GROUP SODIUM? (all other vars are categorical)

```{r, child = "../src/pred_id.Rmd"}

```

\clearpage

## Outcomes 

The following outcomes are considered: 

- CV mortality + first HF hospitalization (is this the primary outcome???)
- CV mortality
- All-cause mortality
- First HF hospitalization
- Total HF hospitalization

Data were censored at 2019-12-31 or death/move from Stockholm county. DO YOU WANT TO CUT OUTCOMES AT A CERTAIN TIMEPOINT?

### Association between ID/A and outcomes

Time to first events were presented with 1-Kaplan-Meier curves for the outcome all-cause mortality 
and cumulative incidence curves treating death as a competing event for other outcomes. 
The mean cumulative function (MCF) was used to depict the recurrent events. 

Cox proportional hazards regressions were 
used to model the time to first event, partly crude including only ID/A in the model 
and partly adjusted for variables indicated in 
Table \ref{tab:tab1aid}. The variables were selected based on clinical relevance. 

As a consistency analysis, the primary outcome CV mortality + first HF hospitalization, 
was modelled using a sub-distributional hazards model [@fg] where death was treated as a 
competing event. 

Recurrent events were model using a negative binomial model 
including the log of time as an offset in the model, crude and adjusted as above. 

The median (min-max) follow-up is 
`r rsdata %>% summarise(med = fn(median(sos_outtime_death / 365.25 * 12), dig = 1),
                                             min = fn(min(sos_outtime_death / 365.25 * 12), dig = 1),
                                             max = fn(max(sos_outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r rsdata %>% 
                                   summarise(sumpy = fn(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.

```{r, child = "../src/km_aid.Rmd"}

```

```{r, child = "../src/outtab_aid.Rmd"}

```

\newpage

### Association between ID and outcomes by EF

Cox proportional hazards regressions were 
used to model the time to first event, partly crude, including only ID in the model 
and partly adjusted for variables indicated in 
Table \ref{tab:tab1aid} partly in the overall cohort and partly 
presented in EF subgroups using interaction effects. 

Recurrent events were model using a negative binomial model 
including the log of time as an offset in the model, crude and adjusted as above. 

```{r, child = "../src/outtab_idef.Rmd"}

```

\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/xx. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
