---
title: 'Statistical report: Phenotyping heart failure patients with iron deficiency'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
urlcolor: blue
linkcolor: black
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

SCREAM2

## Inclusion/exclusion criteria

```{r flow}
footnote(default_kable(flow, caption = "Flowchart"),
  symbol = c("If a patient does not have any posts with hb, ferritin, transferrin at index the first post is selected. If a patient has a first post without lab but a second with lab, the second post is selected.")
)
```

First patient in: `r min(rsdata$shf_indexdtm)` and last patient in: `r max(rsdata$shf_indexdtm)`.  

The median age (IQR) is `r rsdata %>% summarise(med = fn(median(shf_age), dig = 1),
                                             q1 = fn(quantile(shf_age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(shf_age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r rsdata %>% count(shf_sex) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(shf_sex == "Female") %>%
  pull(perc)`% females.    

## Created variables 

```{r npr}
default_kable(metaout, caption = "Comorbidities and outcomes from VAL/KON")
```

The labvalues from SCREAM are selected within 90 days prior to index in SwedeHF, 
where the one closest to index is chosen. If there are two tests of the same lab 
from the same date the mean is calculated. The ferritin, transferrin and 
hb value must be from the same date, the other labs can be from different dates (within the 90 days).

ddd_ironoralsupp is defined as at least one dispensed prescription with ATC strting 
with B03AA, B03AB, B03AD or B03AE within 5 months prior up to 14 days after index. 

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with some care.

## Missing data

Missing data was imputed with multiple imputation (n = 10) using mice [@mice]. 
Variables included in the model are indicated in 
Table \ref{tab:tab1aid}. The imputation is performed only on patients with hb, transferrin and ferritin at index (Iron population). 

As a consistency analysis FCM is assumed to be No for all registrations <=2016 
and for registrations >= 2017 it is imputed with multiple imputation. 

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

\clearpage

## Testing over time 

The denominator is the full population that are alive and living in Stockholm county on 31 December of the respective calender year. 
The numerator is the number of patients that have at least one  
lab measure of hb or ferritin and transferrin (on same date) in the respective calender year. 
Note that this could be prior to inclusion in SwedeHF. 

```{r, child = "../src/overtime.Rmd"}

```

\clearpage

## Time to first lab test after index

Time to first laboratory test after index were presented with cumulative incidence curves, partly for first hb test, 
partly first transferrin and ferritin (on same date) test, and partly all 3 (on same date). 
Data were censored after 3 years, 2018-12-31 or death/move from Stockholm county. For the stratified analysis on patients with index date <= 2016, they were also censored at 2016-12-31. 

The analysis is performed on the full population.

```{r, child = "../src/timefirsttest.Rmd"}

```

\clearpage

## Iron deficiency after index 

The cumulative iron deficiency is presented within 6 mo, 1, 2 and 3 years. At least 
one lab test within the respective time period is necessary to be in the analysis. 
Data were censored 2018-12-31 or death/move from Stockholm county. 
So a patient with iron deficiency at 7 months will not be included in the 
6 mo analysis but will be included in both the numerator and denominator at 1 and 3 years. 

The analysis is performed on the full population. 

```{r, child = "../src/irondef.Rmd"}

```

\clearpage

## Predictors of ID at index

The association between baseline covariates and ID was investigated using 
logistic regression partly in the overall cohort and partly separately in EF subgroups. 

A p-value for interaction between EF and the respective variable (with the other variables in the model) is also presented. 
(I did not present the OR ect from the interaction model because it seems 
a bit untransparant when I thought about it, then the ORs would all be from separate models...). 

The analysis is performed on patients with hb, transferrin and ferritin at index (Iron population). 

```{r, child = "../src/pred_id.Rmd"}

```

\clearpage

```{r, child = "../src/pred_id_fcm.Rmd"}

```

\clearpage

## Outcomes 

The following outcomes are considered: 

- CV mortality + first HF hospitalization
- CV mortality
- All-cause mortality
- First HF hospitalization
- Total HF hospitalization

Data were censored at 3 years after index, 2019-12-31 or death/move from Stockholm county. 

The analysis is performed on patients with hb, transferrin and ferritin at index (Iron population). 

### Association between ID/A and outcomes

Time to first events were presented with cumulative incidence curves for the outcomes. 
The mean cumulative function (MCF) was used to depict the recurrent events. 

Cox proportional hazards regressions were 
used to model the time to first event, partly crude including only ID/A in the model 
and partly adjusted for variables indicated in 
Table \ref{tab:tab1aid}. The variables were selected based on clinical relevance. 

As a consistency analysis, the primary outcome CV mortality + first HF hospitalization, 
was modelled using a sub-distributional hazards model [@fg] where death was treated as a 
competing event. 

Recurrent events were model using a negative binomial model 
including the log of time as an offset in the model, crude and adjusted as above. 

The median (min-max) follow-up is 
`r rsdata %>% summarise(med = fn(median(sos_outtime_death / 365.25 * 12), dig = 1),
                                             min = fn(min(sos_outtime_death / 365.25 * 12), dig = 1),
                                             max = fn(max(sos_outtime_death / 365.25 * 12), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", min, "-", max, ")")) %>%
                                   pull(out)` months for a total of 
                                   `r rsdata %>% 
                                   summarise(sumpy = fn(sum(sos_outtime_death) / 365.25, dig = 0)) %>%
                                   pull(sumpy)` patient-years of follow-up.

```{r, child = "../src/km_aid.Rmd"}

```

\clearpage

```{r, child = "../src/outtab_aid.Rmd"}

```

\clearpage

```{r, child = "../src/outtab_aid_fcm.Rmd"}

```

\clearpage

### Association between ID and outcomes by EF

Cox proportional hazards regressions were 
used to model the time to first event, partly crude, including only ID in the model 
and partly adjusted for variables indicated in 
Table \ref{tab:tab1aid} partly in the overall cohort and partly 
presented in EF subgroups using interaction effects. 

Recurrent events were model using a negative binomial model 
including the log of time as an offset in the model, crude and adjusted as above. 

```{r, child = "../src/outtab_id.Rmd"}

```

```{r, child = "../src/outtab_idef.Rmd"}

```

\clearpage

```{r, child = "../src/outtab_id_fcm.Rmd"}

```

```{r, child = "../src/outtab_idef_fcm.Rmd"}

```

\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/screamiron. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
